---
title: "Sales_Analysis"
author: "Yeji Sohn"
date: '2023-08-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
listOfPackages <- c("dplyr","tidyverse","kableExtra", "tidyr", "broom", "knitr", "ggplot2", "lubridate", "scales", "cowplot")
for (i in listOfPackages){
     if(! i %in% installed.packages()){
         install.packages(i, dependencies = TRUE)
     }
     require(i)
}
```

# Valid value check
```{r}
# Clean data
df = read.csv('data_/online_retail.csv')

# add Date and revenue
df <- df %>% 
  filter(!StockCode == "AMAZONFEE") %>%
  mutate(Date = as.Date(InvoiceDate, "%m/%d/%Y")) %>%
  mutate(revenue  = UnitPrice * Quantity) %>%
  mutate(month = month(Date),
         day = day(Date),
         after15 = ifelse(day > 15, 1, 0),
         month2 = month + 0.5*after15)
```

```{r}
pos_df <- df %>%
  filter(Quantity >= 0 & UnitPrice >= 0)
colSums(df<0)

df %>%
  filter(UnitPrice < 0)
```

Orders with negative quantites are refunds, and orders with negative price are bad debt adjustment. Rows with Negative price will be dropped. 

# Overview 

```{r}
#only purchase
df_pos <- df %>%
  filter(Quantity >= 0)

#only return
df_neg <- df %>%
  filter(Quantity < 0)
```

## number of unique order & return
```{r}
df %>% 
  summarise(number_of_order = length(unique(ifelse(Quantity > 0, InvoiceNo, 0))),
            number_of_return = length(unique(ifelse(Quantity < 0, InvoiceNo, 0)))) %>% 
  mutate(returns_share = number_of_return / number_of_order)

#by items
df %>% 
  summarise(number_of_sold_items = sum(ifelse(Quantity > 0, Quantity, 0)),
            number_of_returned_items = sum(ifelse(Quantity < 0, -Quantity, 0))) %>% 
  mutate(returns_share_items = number_of_returned_items / number_of_sold_items)
#by revenue

df %>% 
  summarise(total_revenue = sum(ifelse(Quantity > 0, Quantity, 0) * UnitPrice, na.rm = TRUE),
            total_returned_revenue = sum(ifelse(Quantity < 0, -Quantity, 0) * UnitPrice, na.rm = TRUE)) %>% 
  mutate(returns_share = total_returned_revenue / total_revenue)

#avg order revenue

df %>% 
  summarise(number_of_unique_customer = length(unique(CustomerID)),
            total_revenue = sum(revenue)) %>% 
  mutate(Avg_revenue_per_order = total_revenue / number_of_unique_customer)


```

```{r}
df2 <- df %>%
  filter(Quantity > 0) %>%
  group_by(Date) %>%
  summarise(order_num  = length(unique(InvoiceNo))) %>% 
  arrange(desc(order_num)) %>%
  ungroup() %>%
  mutate(order_num)

df2 <- df2 %>%
  mutate(month = month(Date),
         day = day(Date),
         after15 = ifelse(day > 15, 1, 0),
         month2 = month + 0.5*after15
         )
```

```{r}
#Number of Order (no return) per month
plot(df2$Date, df2$order_num, xlab = 'Month', ylab = 'Number of Orders')

ggplot(df2, aes(x=month, y=order_num)) + 
  geom_histogram(stat = "identity", fill = 'blue', color = 'white') +
  theme_bw() + 
  scale_x_continuous(expand = c(0,0),breaks=seq(0, 12, by = 1)) + 
  scale_y_continuous(expand = expansion(mult = c(0,0.1))) + 
  labs(x = 'Month',
       y = 'Number of Orders')

#per 0.5 month 
ggplot(df2, aes(x=month2, y=order_num)) + 
  geom_histogram(stat = "identity", fill = 'blue', color = 'white') +
  theme_bw() + 
  scale_x_continuous(expand = c(0,0),breaks=seq(0, 12, by = 1)) + 
  scale_y_continuous(expand = expansion(mult = c(0,0.1))) + 
  labs(x = 'Month',
       y = 'Number of Orders')
```

```{r}
df3 <- df %>%
  filter(Quantity < 0) %>%
  group_by(Date) %>%
  summarise(order_num  = length(unique(InvoiceNo))) %>% 
  arrange(desc(order_num)) %>%
  ungroup() %>%
  mutate(order_num)

df3 <- df3 %>%
  mutate(month = month(df3$Date),
         day = day(df3$Date),
         after15 = ifelse(day > 15, 1, 0),
         month2 = month + 0.5*after15
         )

```

```{r}
#Number of Returns per month
plot(df3$Date, df3$order_num, xlab = 'Month', ylab = 'Number of Returns')

ggplot(df3, aes(x=month, y=order_num)) + 
  geom_histogram(stat = "identity", fill = 'blue', color = 'white') +
  theme_bw() + 
  scale_x_continuous(expand = c(0,0),breaks=seq(0, 12, by = 1)) + 
  scale_y_continuous(expand = expansion(mult = c(0,0.1))) + 
  labs(x = 'Month',
       y = 'Number of Returns')

#per 0.5month
ggplot(df3, aes(x=month2, y=order_num)) + 
  geom_histogram(stat = "identity", fill = 'blue', color = 'white') +
  theme_bw() + 
  scale_x_continuous(expand = c(0,0),breaks=seq(0, 12, by = 1)) + 
  scale_y_continuous(expand = expansion(mult = c(0,0.1))) + 
  labs(x = 'Month',
       y = 'Number of Returns')
```

```{r}
#revenue per month
df4 <- df %>%
  filter(Quantity > 0) %>%
  group_by(month) %>%
  summarise(month_revenue  = sum(revenue)) %>% 
  mutate(month_revenue)

p <- ggplot(df4, aes(x=month, y=month_revenue)) + 
  geom_histogram(stat = "identity", fill = 'blue', color = 'white') + 
  theme_bw() + 
  scale_x_continuous(expand = c(0,0), breaks=seq(0, 12, by = 1)) +
  scale_y_continuous(expand = expansion(mult = c(0,0))) +
  labs(x = 'Month',
       y = 'Revenue')

point <- format_format(big.mark = " ", decimal.mark = ",", scientific = FALSE)

p + scale_y_continuous(labels = point)

#per0.5 month 
df4b <- df %>%
  filter(Quantity > 0) %>%
  group_by(month2) %>%
  summarise(month_revenue  = sum(revenue)) %>% 
  mutate(month_revenue)

p <- ggplot(df4b, aes(x=month2, y=month_revenue)) + 
  geom_histogram(stat = "identity", fill = 'blue', color = 'white') + 
  theme_bw() + 
  scale_x_continuous(expand = c(0,0), breaks=seq(0, 12, by = 1)) +
  scale_y_continuous(expand = expansion(mult = c(0,0))) +
  labs(x = 'Month',
       y = 'Revenue')

p + scale_y_continuous(labels = point)

##################################
#filter the biggest 2 order that have been returned
df4 <- df %>%
  filter(Quantity > 0) %>%
  filter(!StockCode == "23843" & !StockCode == "23166") %>%
  group_by(month) %>%
  summarise(month_revenue  = sum(revenue)) %>% 
  mutate(month_revenue)

p <- ggplot(df4, aes(x=month, y=month_revenue)) + 
  geom_histogram(stat = "identity", fill = 'black', color = 'white') + 
  theme_bw() + 
  scale_x_continuous(expand = c(0,0), breaks=seq(0, 12, by = 1)) +
  labs(x = 'Month',
       y = 'Revenue')

p + scale_y_continuous(labels = point)

```

```{r}
df5 <- df %>%
  filter(Quantity < 0) %>%
  group_by(month) %>%
  summarise(return_amount  = abs(sum(revenue))) %>% 
  mutate(return_amount)

p <- ggplot(df5, aes( x = month, y = return_amount)) + 
  geom_histogram(stat = "identity", fill = 'blue', color = 'white') +
  theme_bw() + 
  scale_x_continuous(expand = c(0,0),breaks=seq(0, 12, by = 1)) + 
  scale_y_continuous(expand = expansion(mult = c(0,0.1))) + 
  labs(x = 'Month',
       y = 'Return Amount')

p + scale_y_continuous(labels = point)

#per 0.5 month

df5b <- df %>%
  filter(Quantity < 0) %>%
  group_by(month2) %>%
  summarise(return_amount  = abs(sum(revenue))) %>% 
  mutate(return_amount)

p <- ggplot(df5b, aes( x = month2, y = return_amount)) + 
  geom_histogram(stat = "identity", fill = 'blue', color = 'white') +
  theme_bw() + 
  scale_x_continuous(expand = c(0,0),breaks=seq(0, 12, by = 1)) + 
  scale_y_continuous(expand = expansion(mult = c(0,0.1))) + 
  labs(x = 'Month',
       y = 'Return Amount')

p + scale_y_continuous(labels = point)

#############################
#filter the biggest 2 order that have been returned
df5 <- df %>%
  filter(Quantity < 0) %>%
  filter(!StockCode == "23843" & !StockCode == "23166") %>%
  group_by(month) %>%
  summarise(return_amount  = abs(sum(revenue))) %>% 
  mutate(return_amount)

p <- ggplot(df5, aes( x = month, y = return_amount)) + 
  geom_histogram(stat = "identity", fill = 'black', color = 'white') + 
  theme_bw() + 
  scale_x_continuous(breaks=seq(0, 12, by = 1)) + 
  labs(x = 'Month',
       y = 'Return Amount')

p + scale_y_continuous(labels = point)

#per 0.5 month

df5b <- df %>%
  filter(Quantity < 0) %>%
  filter(!StockCode == "23843" & !StockCode == "23166") %>%
  group_by(month2) %>%
  summarise(return_amount  = abs(sum(revenue))) %>% 
  mutate(return_amount)

p <- ggplot(df5b, aes( x = month2, y = return_amount)) + 
  geom_histogram(stat = "identity", fill = 'blue', color = 'white') +
  theme_bw() + 
  scale_x_continuous(expand = c(0,0),breaks=seq(0, 12, by = 1)) + 
  scale_y_continuous(expand = expansion(mult = c(0,0.1))) + 
  labs(x = 'Month',
       y = 'Return Amount')

p + scale_y_continuous(labels = point)

```

## Characteristics of customers who got discount
```{r}
df <- read.csv("data_/online_retail.csv")
d <- as.POSIXct(df$InvoiceDate, format = "%m/%d/%Y %H:%M")
df$InvoiceDate <- d

discount <- df %>%
  filter(StockCode == "D")  %>%
  select(CustomerID)
d <- df %>%
  inner_join(discount, by = "CustomerID")
sample_n(d, 10)
```

## Transaction volume variation over day of time
```{r}
# If InvoiceDate is in local time
temp <- df %>% 
  mutate(Hour = as.integer(strftime(InvoiceDate, format="%H")))
min(temp$Hour)
max(temp$Hour)
```

```{r}
hist(temp$Hour)
```

Most orders are made middle of the date. Note that there is no orders before 6 am and after 9 pm. The distribution is approximately normal.

# Understanding Returns

**Returned products are represented as negative values in the dataset**

```{r}
df <- read_csv("data_/online_retail.csv")
df$revenue <- df$Quantity * df$UnitPrice 
returned_df = df[df$Quantity < 0, ]
retProds <- returned_df %>% 
  group_by(StockCode, Description, UnitPrice, CustomerID, quantRet = abs(Quantity)) %>% #quantRet = quantity returned
  summarise(revenue_returned = sum(abs(revenue)))

top_retProds_noAMZN <- head(retProds[order(-retProds$revenue_returned), ] %>% filter(StockCode != 'AMAZONFEE'), 10)
kable(top_retProds_noAMZN) %>% kable_styling(latex_options = "striped")
```
The products with the most quantity and revenue returned are _Paper Craft , LITTLE BIRDIE_ and _MEDIUM CERMAIC TOP STORAGE JAR_. However, there's still some uncertainty shrouding _AMAZON FEE_ as it accounts for bottom 7 of top 10 list of products with the most returns, but it is registered as a distinct products instead of 1. Also there's no CustomerID so it's impossible to know who is purchasing this _AMAZON FEEs_.

```{r}
retProdsLean <- returned_df %>% 
  group_by(StockCode, Description) %>% #grouping amazon fee and manual 
  summarise(quantity_returned = sum(abs(Quantity)),revenue_returned = sum(abs(revenue)))
top_retProds_noAMZNLean <- head(retProdsLean[order(-retProdsLean$revenue_returned), ], 10)
kable(top_retProds_noAMZNLean) %>% kable_styling(latex_options = "striped")
#table(returned_df[order(-returned_df$revenue_returned), ]$Description )
```

# Understanding Market

```{r}
df = read.csv('data_/online_retail.csv')
d = as.integer(format(as.POSIXct(df$InvoiceDate, format = "%m/%d/%Y %H:%M"), "%Y"))
df$InvoiceDate <- d
names(df)[names(df) == 'InvoiceDate'] <- 'year'
gdp = read.csv('data_/gdp_per_capita.csv')
gdp = gdp %>% select("Country.Name", "Code", "X2011", "X2010")
```

```{r}
df_pc = left_join(df, gdp, by=c("Country" = "Country.Name"))

df_pc = df_pc %>%
  filter(Quantity > 0) %>%
  filter(UnitPrice > 0)

df_pc = df_pc %>%
  mutate(gdp_pc = ifelse(year == 2010, X2010, X2011))  %>%
  mutate(revenue = UnitPrice*Quantity)

df_pc_rev = df_pc %>%
  group_by(Country, year) %>%
  summarize(revenue = sum(revenue))

df_gdp = df_pc %>% select(Country, year, gdp_pc, Code)  %>% distinct
df_viz = left_join(df_pc_rev, df_gdp, by=c("Country", "year"))
```